---
title: "interactive-contact-conversion"
author: "Dan Spakowicz"
date: "12/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Read in data

Export contacts as **CSV** (Google Contacts or compatible; column names must match what this script expects).

This script reads a CSV from `../data/g-contacts_exports/`. Update the path below when you use a new export.

```{r read-google-contacts, message=FALSE}
contacts_path <- "../data/g-contacts_exports/2025-11-11_contacts.csv"

contacts_raw <- read_csv(contacts_path)

head(contacts_raw)
```

## Select and clean columns

```{r select-columns}
contacts <- contacts_raw %>%
  select(
    `First Name`,
    `Last Name`,
    `Home Street`,
    `Home City`,
    `Home State`,
    `Home Postal Code`
  ) %>%
  drop_na(`Home City`)

head(contacts)
```

## Merge duplicate addresses

Contacts at the same address are merged into one row. Matching uses the first two words of `Home Street` (e.g. house number + street name) so small wording differences (Rd vs Road) still group together.

```{r merge-duplicates}
contacts <- contacts %>%
  mutate(street_key = word(`Home Street`, 1, 2)) %>%
  group_by(street_key) %>%
  summarise(
    `First Name` = if_else(
      n() > 1,
      paste(`First Name`, collapse = " & "),
      first(`First Name`)
    ),
    `Last Name` = if_else(
      n_distinct(`Last Name`) > 1,
      paste(unique(`Last Name`), collapse = " & "),
      first(`Last Name`)
    ),
    `Home Street` = first(`Home Street`),
    `Home City` = first(`Home City`),
    `Home State` = first(`Home State`),
    `Home Postal Code` = first(`Home Postal Code`),
    .groups = "drop"
  ) %>%
  select(-street_key)
```

## Export outputs

All exports go to `../data/script-outputs/` with a run date in the filename.

```{r export-date}
export_date <- Sys.Date()
```

### SimplyToImpress

Column names stay aligned with the Google export.

```{r export-simplytoimpress}
write.csv(
  contacts,
  file = paste0("../data/script-outputs/", export_date, "_holiday_simplytoimpress.csv"),
  row.names = FALSE,
  na = ""
)
```

### Shutterfly

Shutterfly expects different column names (no `Home ` prefix on address fields, plus extra empty columns).

```{r export-shutterfly}
remove_home_prefix <- function(x) gsub("Home (.*)", "\\1", x)

shutterfly <- contacts %>%
  mutate(
    `Display/Household Name (Smith Family)` = "",
    `Address Line 2` = "",
    `Country` = "",
    `Email` = ""
  ) %>%
  rename_with(remove_home_prefix, c("Home City", "Home State", "Home Postal Code")) %>%
  rename(`Address Line 1` = `Home Street`) %>%
  select(
    `First Name`,
    `Last Name`,
    `Display/Household Name (Smith Family)`,
    `Address Line 1`,
    `Address Line 2`,
    `City`,
    `State`,
    `Postal Code`,
    `Country`,
    `Email`
  )

head(shutterfly)

write.csv(
  shutterfly,
  file = paste0("../data/script-outputs/", export_date, "_holiday_shutterfly.csv"),
  row.names = FALSE,
  na = ""
)
```

### Paper Culture

Full state names are required; `state.name` / `state.abb` map two-letter codes.

```{r export-paperculture}
paper_culture <- contacts %>%
  mutate(
    `Recipient Name (enter 1 per line)` = paste(`First Name`, `Last Name`),
    USAddress1 = `Home Street`,
    USAddress2 = NA,
    USCity = `Home City`,
    USState = state.name[match(`Home State`, state.abb)],
    USZipCode = `Home Postal Code`
  ) %>%
  select(
    `Recipient Name (enter 1 per line)`,
    USAddress1,
    USAddress2,
    USCity,
    USState,
    USZipCode
  )

write.csv(
  paper_culture,
  file = paste0("../data/script-outputs/", export_date, "_holiday_paperculture.csv"),
  row.names = FALSE,
  na = ""
)
```
