---
title: "interactive-contact-conversion"
author: "Dan Spakowicz"
date: "12/21/2020"
output: html_document
---

```{r setup, include=FALSE}
# Set knitr options to display code chunks
knitr::opts_chunk$set(echo = TRUE)

# Load required packages
library(tidyverse)
```

## Read in data

Data must be exported in OUTLOOK format

This script reads a CSV file containing exported Google Contacts data.
The file path provided points to a specific export dated "2024-12-14".
Ensure the file exists at the specified relative path before running the script.

```{r read in google contacts export, message=FALSE}
# Read the Google Contacts export CSV file
# Note: Update the file path/date as needed for your export
x <- read_csv("../data/g-contacts_exports/2025-11-11_contacts.csv")

# Preview the data structure
head(x)
```

## Select and clean columns

```{r}
# Select only the columns needed for holiday card addressing
# Remove any contacts that don't have a city (i.e., incomplete addresses)
df <- x %>%
  select(`First Name`, `Last Name`, `Home Street`,
         `Home City`, `Home State`, `Home Postal Code`) %>%
  drop_na(`Home City`)

# Preview the cleaned data
head(df)
```

## Merge duplicate addresses

```{r}
# Merge contacts that share the same address
# Strategy: Extract first two words of street address (number + street name)
# to match addresses that might differ only in abbreviations (Rd vs Road, etc.)

df <- df %>%
  # Create a simplified address identifier (e.g., "123 Main")
  mutate(num.st = word(`Home Street`, 1, 2)) %>%
  # Group by this simplified address
  group_by(num.st) %>%
  summarise(
    # If multiple people at same address, combine first names with " & "
    `First Name` = if_else(n() > 1, paste(`First Name`, collapse = " & "), first(`First Name`)),
    # If different last names at same address, combine them with " & "
    `Last Name` = if_else(n_distinct(`Last Name`) > 1, 
                          paste(unique(`Last Name`), collapse = " & "), 
                          first(`Last Name`)),
    # Use the first occurrence of address fields (they should be identical)
    `Home Street` = first(`Home Street`),
    `Home City` = first(`Home City`),
    `Home State` = first(`Home State`),
    `Home Postal Code` = first(`Home Postal Code`),
    .groups = "drop"
  ) %>%
  # Remove the temporary grouping variable
  select(-num.st)
```

## Export for SimplyToImpress

```{r}
# Generate date stamp for output filename
date <- Sys.Date()

# Export cleaned data in SimplyToImpress format
# This format uses the original column names from Google Contacts
write.csv(df, 
          file = paste0("../data/script-outputs/", date, "_holiday_simplytoimpress.csv"),
          row.names = FALSE,  # Don't include row numbers
          na = "")            # Write empty string instead of NA
```

## Export for Shutterfly

```{r}
# Helper function to remove "Home " prefix from column names
# Shutterfly expects "City" not "Home City", etc.
remove_home_prefix <- function(x) gsub("Home (.*)", "\\1", x)

# Transform data to match Shutterfly's required format
shut <- df %>%
  # Add empty columns required by Shutterfly
  mutate("Display/Household Name (Smith Family)" = "",
         "Address Line 2" = "",
         "Country" = "", 
         "Email" = "") %>%
  # Remove "Home " prefix from city, state, and postal code columns
  rename_with(remove_home_prefix, c("Home City", "Home State", "Home Postal Code")) %>%
  # Rename street address to match Shutterfly format
  rename("Address Line 1" = "Home Street") %>%
  # Select columns in the exact order Shutterfly expects
  select("First Name", "Last Name", "Display/Household Name (Smith Family)", 
         "Address Line 1", "Address Line 2", "City", "State", "Postal Code", 
         "Country", "Email")

# Preview the transformed data
head(shut)

# Export to CSV file
write.csv(shut, 
          file = paste0("../data/script-outputs/", date, "_holiday_shutterfly.csv"),
          row.names = FALSE, 
          na = "")
```

## Export for Paper Culture

```{r}
# Transform data to match Paper Culture's required format
pc <- df %>%
  mutate(
    # Combine first and last name into a single recipient name field
    "Recipient Name (enter 1 per line)" = paste(`First Name`, `Last Name`),
    # Map address fields to Paper Culture's naming convention
    USAddress1 = `Home Street`,
    USAddress2 = NA,  # No second address line in our data
    USCity = `Home City`,
    # Convert state abbreviations (e.g., "CA") to full state names (e.g., "California")
    USState = state.name[match(`Home State`, state.abb)],
    USZipCode = `Home Postal Code`
  ) %>%
  # Select only the columns Paper Culture requires, in order
  select(`Recipient Name (enter 1 per line)`, 
         USAddress1, 
         USAddress2, 
         USCity, 
         USState, 
         USZipCode)

# Export to CSV file
write.csv(pc, 
          file = paste0("../data/script-outputs/", date, "_holiday_paperculture.csv"),
          row.names = FALSE, 
          na = "")
```
